Probabilistic Policy Violation Detection with IPFWD

### 2017-01-23 paper/main.tex
Add LaTeX paper outline

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{IEEE Journal Template Example}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
        John~Doe,~\IEEEmembership{Fellow,~OSA,}
        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\thanks{M. Shell is with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

@@

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

@@

\begin{abstract}
%\boldmath
\blindtext[1]
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

@@

\section{Introduction}
\blindtext

\subsection{Subsection Heading Here}
\blindtext

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\section{Conclusion}
\blindtext

@@

@@

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Probabilistic Policy Violation Detection with IPFWD}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
\thanks{placeholder}}

@@

% The paper headers
\markboth{March~2017}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

\begin{abstract}
%\boldmath
When using a firewall system like IPFW to detect threats, we can end up doing a lot of packet processing. This can negatively impact performance-sensitive systems such as storage nodes. This paper describes a practical solution to this problem using a load-weighted probabilistic mechanism that allows a trade-off between perfect visibility of incoming packets and reduced impact to system load.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
FreeBSD, IPFW, firewalls, performance.
\end{IEEEkeywords}

@@

\section{Introduction}

IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a chance to accept any packet. The probability of acceptance is updated over time and dependent on the current system load. To account for this, IPFWD extends IPFW logs to include the likelihood an undetected policy violation occurred.

This approach is based on the premise that firewall performance can be improved by reducing the number of rules applied to each packet. Supposing a whitelist policy, a firewall has to apply every rule to each packet before it's denied. With IPFWD, we have a chance to accept any packet early and skip any further computation. Test results shows this reduces the system resources required to handle the same amount of traffic.

This is a shift in mindset from typical firewalls. Instead of enforcing every part of a security policy all the time, IPFWD enforces the policy some of the time and extrapolates from the violations it encounters. This provides insight into the violations that weren't detected. For this cost, you gain increased firewall performance and network throughput in resource bound systems. It's acceptable to allow a percentage of policy violations given that network traffic patterns are often repeated and the goal is detection and not immediate prevention. Preventative action may be taken later when resource requirements are lower.

As an example, under heavy load IPFWD may immediately accept 40\% of packets. Supposing a port scan was initiated during this time and rules exist to block it, 60\% of the port scan would still be rejected and logged. Since the probability will fluctuate over time, IPFWD provides information in the IPFW logs to show the chance undetected violations occurred for each detected violation. IPFWD allows administrators to keep extensive rule sets that fully implement their security policy. Instead of having to simplify rule sets to increase performance, IPFWD balances policy enforcement and performance automatically. Under normal or light load, IPFWD will enforce the entire security policy 100\% of the time.

\section{IPFW}
% Why was this firewall chosen?
% First match wins rule system
% Built in probabilistic match
\blindtext

\section{IPFWD}
% What does it do?
% How does it work?
% - Probabilistic packet matching
% - Log extension
% Explain rationale for it acting as it does
% When should this system be used?
% - High performance
% - Resource conscious
\blindtext

\section{Security Implications}
% Talk about security policy enforcement
% What does it mean for security if your ignoring some of the packet?
% How do you trust this system?
% Talk about shift in mindset from "blocking attacks" to "detecting attacks"
% - Why is this reasonable?
% - Is this really different than traditional firewalls?
\blindtext

\section{Benchmarking Performance}
% Explain different systems
% Explain tools
% Explain rationale for choosing tools
\blindtext

\section{Results}
% Show results
% Difference in results between CPU and throughput bound systems
\blindtext

\section{Related Work}
% Cisco and Juniper routers
% Machine learning firewalls
\blindtext

\section{Future Work}
% Firewall rule reordering so we can "fail as early as possible"
\blindtext

\section{Conclusion}
% What performance benefit do we get?
% What security cost is there?
% Why is this useful?
% Why is this important?
\blindtext

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{picture}}]{John Doe}
\blindtext
\end{IEEEbiography}

@@

% NOTES
%==========================================================
% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol
